---
title: "ComplexHeatmap"
date: "23/8/2019"
output: html_document
---

### Install ComplexHeatmap with devtools

### This command is only necessary when the script is run for the first time. If the package is installed, go to the following command "load the library"

```{r}
library(devtools)
install_github("jokergoo/ComplexHeatmap")
```

### Load the library
```{r}
library(ComplexHeatmap)
```

### Load variants saved either in csv or txt
```{r}
ToHeatmap <- write.csv("VariantstoHeatmap.csvc", sep = ",", header = TRUE)
head(ToHeatmap)
```

### With "sort" select the variables that want to include in the heatmap
### Our data have three columns (first column - Sample; second column - Gene, and third - Type)
```{r}

nameValsSample <- sort(unique(unlist(ToHeatmap[,1]))) # First column
nameValsGene <- sort(unique(unlist(ToHeatmap[,2]))) # Second column

# construct matrix of correct dimensions with row and column names
myMat <- matrix("", length(nameValsSample),
                length(nameValsGene), dimnames = list(nameValsSample, nameValsGene))

# fill in the matrix with matrix indexing on row and column names
myMat[as.matrix(ToHeatmap[c("Sample",
                             "Gene")])] <- as.character(ToHeatmap[["Type"]])   

```


```{r}
# with the functions below we set the heatmap colors
# gp = gpar(fill = "#F8FCFC" ... ) - it is the color of the variables in the oncoprint
# gp = gpar(..., col = "black") - is the color of the contour

get_type_fun_onoco = function(x) strsplit(x, ";")[[1]]

alter_fun = list(
  background = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#F8FCFC", col = "black"))
  },
  stopgain = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#721210", col = "black"))
  },
  Hotspot = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#ED903D", col = "black"))
  },
  Gain = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#FF0000", col = "black"))
  },
  Loss = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#005EFF", col = "black"))
  },
  Frameshift_deletion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#084A61", col = "black"))
  },
  Nonsynonymous_SNV = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#85B013", col = "black"))
  },
  Frameshift_insertion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(0.1, "mm"), h-unit(0.1, "mm"), gp = gpar(fill = "#13B0A3", col = "black"))
  }
)

col = c("stopgain"="#721210",
        "Hotspot"="#ED903D",
        "Gain"="#FF0000",
        "Loss"="#005EFF",
        "Frameshift_deletion"="#084A61",
        "Nonsynonymous_SNV"="#85B013",
        "Frameshift_insertion"="#13B0A3")
```


### with this function plot the heatmap
```{r}
ht <- oncoPrint(t(myMat), alter_fun = alter_fun, col = col, show_column_names = TRUE,
                left_annotation =  rowAnnotation(rbar = anno_oncoprint_barplot(
                    axis_param = list(side = "top", direction = "reverse"), 
                    border = FALSE, width = unit(1, "cm"))), 
                right_annotation = NULL,
                top_annotation = HeatmapAnnotation(cbar = anno_oncoprint_barplot(border = FALSE, 
                                                          height = unit(1, "cm"))))


draw(ht)
```

